---
title: "LDA"
author: "MADON KENGNE Sirianne"
date: "2022-12-19"
output: html_document
---


## Topic:

The goal is to propose a method for leaf identification based on the provided leaf attributes and using a proper unsupervised or supervised learning tool.

```{r}
#install.packages("tidyverse")
library(tidyverse)
```

```{r}
#install.packages("mlr", dependencies = TRUE)
library(mlr)
```

```{r}
#library(ggplot2)
#library(cowplot)
#library(randomForest)
```


Let's read the data and see the first 6 rows.
```{r}
data <- read.csv("leaf.csv", header=FALSE)
#head(data)
```
Let's give proper columns names to our data
```{r}
colnames(data) <- c("Species","Specimen Number","Eccentricity","AspectRatio","Elongation","Solidity","Stochastic_Convexity","Isoperimetric_Factor","Maximal_Indentation_Depth","Lobedness","Average_Intensity","Average_Contrast","Smoothness","Third_moment","Uniformity","Entropy")

#names(data)
```
Let's give the proper name of species
```{r}
last_species_names<-c(1:15,22:36)
new_species_names<-c("Quercus suber","Salix atrocinera","Populus nigra","Alnus sp.","Quercus robur",
     "Crataegus monogyna","Ilex aquifolium","Nerium oleander","Betula pubescens",
     "Tilia tomentosa","Acer palmatum","Celtis sp.","Corylus avellana","Castanea sativa","Populus alba",16:21,
     "Primula vulgaris","Erodium sp.","Bougainvillea sp.","Arisarum vulgare","Euonymus japonicus","Ilex perado ssp. azorica",
     "Magnolia soulangeana","Buxus sempervirens","Urtica dioica","Podocarpus sp.","Acca sellowiana","Hydrangea sp.","Pseudosasa japonica",
     "Magnolia grandiflora","Geranium sp.")
for(i in last_species_names){
  data[data$Species == i,]$Species <-new_species_names[i] 
} 

#unique(data$Species)
```

let's delete the columns 2 because it is useless for what we want to do
```{r}
dat<-data[,-2]
#head(dat)
```


```{r}
dat$Species <- as.factor(dat$Species)
#str(dat)
```


```{r}
#summary(dat)
```

we will check the number available for each species
```{r}
#table(dat$Species)
```

Let partition our data by creating two independent samples.
```{r}
set.seed(123)
ind<-sample(2, nrow(dat), replace=TRUE, prob=c(0.7,0.3))
training<-dat[ind==1,]
testing<-dat[ind==2,]
```


```{r}
#create a learner
#set.seed(12345)
#dataTask <- makeClassifTask(data = train, target = "Species")
#qda <- makeLearner("classif.qda")
#qdaModel <- train(qda, dataTask)
```

```{r}
library(klaR)
library(psych)
library(MASS)
```

# Linear Discriminant Analysis (LDA) 

```{r}
#https://www.r-bloggers.com/2021/05/linear-discriminant-analysis-in-r/
linear <- lda(Species~., training)
linear

```

```{r}
#prediction
p <- predict(linear, testing)
p1 <- predict(linear, training)$class
#tab <- table(Predicted = p1, Actual = training$Species)
#sum(diag(tab))/sum(tab)
```


```{r}
p2 <- predict(linear, testing)$class
tab1 <- table(Predicted = p2, Actual = testing$Species)
#sum(diag(tab1))/sum(tab1)
tab1
```


```{r}
accuracy <- mean(p2== testing$Species)
accuracy
```

```{r}
library(caret)
confusionMatrix(p2, testing$Species, mode = "everything", positive="1") #library(caret)
```

```{r}
#model
library(pROC) #for ROC curve

roc_lda = multiclass.roc(response = testing$Species,
              predictor = p$posterior) #the second col. contains the probabilities for the yes category

auc(roc_lda)
```

```{r}
set.seed(123)
cv_results <- train(Species ~ ., data = train, method = "lda", trControl = trainControl(method = "cv", number = 10))
cv_results
```
